BETTING AND BELIEF: PREDICTION MARKETS AND ATTRIBUTION OF CLIMATE CHANGE

### 2016-03-29 prediction_market_wsc.tex

As a validation that the market is operating correctly, we run the market with actual temperatures 1880 to 2015 and market betting from 1931 to 2014. Before 1980 there is no statistically significant warming trend so we would not expect the market to change beliefs before then. As a global warming signal begins to show up in 1980s, traders begin to converge toward the log(\ce{CO2}) model (Fig.~\ref{fig:time}).

@@

We use the model to perform 10 full simulations for each of the 500 input parameter sets and average the 10 convergence scores. We conducted multiple simulations for the same parameter set because there is stochasticity in the tempertature time series, the social network structure, and the agent decision models. We then conducted a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of mean belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}. Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).  The difference between the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships: every value, $y_s$, in the data set is replaced by $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, 
where $y_r$ are the observed values of $y$ in the data set. We conduct 1,000 bootstrapped estimations of the partial rank correlation coefficients to obtain 95\% confidence intervals.

@@

The sensitivity analysis averages over time and thus masks time trends. We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and conducted an experiment crossing 95\textsuperscript{th} and 5\textsuperscript{th} percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model. We then collected the time series of convergence across these eight designs for visualization of the distributions of convergence over time (Fig.~\ref{fig:time}). Under most parameterizations, within 15--20 years, the median fraction of traders believing in the true model is over 75\%.

@@

What is not clear to us is why increasing the number of traders slows convergence, even if the number of edges per trader remains fixed. This will be a topic for future research where we will be increasing the number of traders into the thousands.

The historical sequence in Fig.~\ref{fig:time} shows convergence to the TSI model until the early 1970s, after which traders converge toward the \ce{CO2} model. This is consistent with the fact that greenhouse gas concentrations did not become high enough to begin dominating natural climate variations until the 1970s or so, and it took until the 1990s for the temperature record to show clear signs of anthropogenic interference \shortcite{IPCC_WG1_AR5_2013}.

>>>>

As a validation that the market is operating correctly, we ran the market using actual historical temperatures from 1880 to 2014, with market betting from 1931 to 2014. Greenhouse gas concentrations did not become high enough to begin dominating natural climate variations until the 1970s or so, and it took until the 1990s for the temperature record to show clear signs of anthropogenic interference \shortcite{IPCC_WG1_AR5_2013}. Consistent with this, our simulated historical trading sequence (Fig.~\ref{fig:time}) shows convergence to belief in the TSI model until the early 1970s, after which traders converge toward believing the \ce{CO2} model. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[t]

@@

We used the model to perform 10 full simulations for each of the 500 input parameter sets and average the 10 convergence scores. We conducted multiple simulations for the same parameter set because there is stochasticity in the tempertature time series, the social network structure, and the agent decision models. We then conducted a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of mean belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}. Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).  The difference between the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first rank-transformed in order to allow us capture potentially non-linear relationships: every value, $y_s$, in the data set is replaced by $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, 
where $y_r$ are the observed values of $y$ in the data set. We conduct 1,000 bootstrapped estimations of the partial rank correlation coefficients to obtain 95\% confidence intervals.

@@

The sensitivity analysis averages over time and thus masks time trends. We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and conducted an experiment crossing 95\textsuperscript{th} and 5\textsuperscript{th} percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model. We collected the time series of belief-convergence across these eight designs to visualize the distributions of convergence over time (Fig.~\ref{fig:time}). Under most parameterizations, the median fraction of traders believing in the true model reaches 75\% in 10--20 years.

@@

What is not clear to us is why increasing the number of traders slows convergence, even if the number of edges per trader remains fixed. This is a topic for future research.
